---
title: "TEA Special Education"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
```

```{r}
# cstud15 download
download.file("https://github.com/utdata/rwd-mastery-assignments/blob/master/ed-special-ed/data/CSTUD15.csv?raw=true", "data-raw/CSTUD15.csv")
# cstud20 download
download.file("https://github.com/utdata/rwd-mastery-assignments/blob/master/ed-special-ed/data/CSTUD20.csv?raw=true", "data-raw/CSTUD20.csv")
# directory download
download.file("https://github.com/utdata/rwd-mastery-assignments/blob/master/ed-special-ed/data/Directory.csv?raw=true", "data-raw/Directory.csv")
```
## Import Data
We are importing the school ratings data
```{r}
cstud15 <- read_csv("data-raw/cstud15.csv") %>% 
  clean_names()
```
```{r}
# peek at the data
cstud15 %>% glimpse()
```
```{r}
cstud20 <- read_csv("data-raw/cstud20.csv") %>% 
  clean_names()
```
```{r}
# peek at the data
cstud20 %>% glimpse()
```
```{r}
directory <- read_csv("data-raw/Directory.csv") %>% 
  clean_names()
```
```{r}
# peek at the data
directory %>% glimpse()
```
```{r}
directory_cols <- directory %>% 
  select(
    school_number,
    school_name,
    district_name,
    instruction_type,
    charter_type,
    grade_range
  ) %>% 
  mutate(
    campus = school_number %>% str_replace("'","")
  ) %>% 
  select(-school_number)
  
directory_cols %>% glimpse()
```
Preparing the student enrollment data
```{r}
# set up 2015
cstud15_cols <- cstud15 %>% 
  select(
    campus,
    cpetspec,
    cpetspep
  ) %>% 
  rename(
    sped_c_15 = cpetspec,
    sped_p_15 = cpetspep
  )

# set up 2020
cstud20_cols <- cstud20 %>% 
  select(
    campus,
    cpetspec,
    cpetspep
  ) %>% 
  rename(
    sped_c_20 = cpetspec,
    sped_p_20 = cpetspep
  )

# peek at both
cstud15_cols %>% glimpse()
```
we need the schools that were open in both 2015 and 2020, but some schools existed in only one year or the other, so we’ll drop those. When we join our files, we’ll use an inner_join() to keep only schools with matching records from both 2015 and 2020. This will drop schools that opened after 2015 or closed before 2020
```{r}
sped <- directory_cols %>% 
  inner_join(cstud15_cols)

## Joining, by = "campus"
sped <- directory_cols %>% 
  inner_join(cstud15_cols) %>% 
  inner_join(cstud20_cols)
sped %>% glimpse()
```
## Exports
exporting files to .rds
```{r}
sped %>%
  write_rds("data-processed/01_sped.rds")
```
