---
title: "Special education computations"
output: html_notebook
---

To explore an analyze our school ratings data. Mention that you have to run the other notebook first in case your someone else (or your future self) comes here first.

## Goals
- Which campus gained the most (count difference) special education students from 2015 to 2020?
- Which campus has the highest share (percent) of special education students in 2020?
- Which campus had the greatest change in share of in special education students between 2015 and 2020.
- How many AISD schools would be above the special education "audit threshold" of 8.5% in 2020 if it were still in effect? How does those numbers compare to 2015?

We will filter our list to focus on "Regular Instructional" schools to exclude alternative, disciplinary or schools part of the justice system.

##Import
```{r setup}
library(tidyverse)
library(janitor)
```
##Loading the Data
```{r import}
sped <- read_rds("data-processed/01_sped.rds")

sped %>% glimpse()
```
##Exploring instruction_type
count the rows based on values.
```{r explore-type}
sped %>% 
  count(instruction_type)
```
## Filter fun
filtering for regular schools
rewrote the first filter to indent it differently so it was more clear that we have two logical tests here. The second test has a nested function, testing the charter_type field to see if it is blank, or is.na(), which is a special case. We could have written this line as is.na(charter_type).
Add a select function and remove the charter_type and instruction_type columns.
Assign the result to a new data frame called sped_reg.
Print the data frame to the screen so you can poke around
```{r}
sped_reg <- sped %>% 
  filter(
    instruction_type == "REGULAR INSTRUCTIONAL",
    charter_type %>% is.na()
  ) %>% 
  select(-charter_type, -instruction_type)

sped_reg
```
```{r}
sped_calcs <- sped_reg %>%
  mutate(
    #simple difference
    sped_c_diff = sped_c_20 - sped_c_15
  )
sped_calcs %>% glimpse()
```
```{r}
sped_calcs %>% 
  select(school_name, sped_c_15, sped_c_20, sped_c_diff) %>% 
  head()
```
```{r}
sped_calcs <- sped_reg %>% 
  mutate(
    # simple diff
    sped_c_diff = sped_c_20 - sped_c_15,
    # percent change
    spec_c_prccng = ((sped_c_20 - sped_c_15) / sped_c_15 * 100) %>% round_half_up(1),
    # simple diff of percent
  )

sped_calcs %>% glimpse()
```
```{r}
sped_calcs <- sped_reg %>% 
  mutate(
    # simple diff
    sped_c_diff = sped_c_20 - sped_c_15,
    # percent change
    sped_c_prccng = ((sped_c_20 - sped_c_15) / sped_c_15 * 100) %>% round_half_up(1),
    # simple diff of percent
    sped_p_ppd = sped_p_20 - sped_p_15
  )

sped_calcs %>% glimpse()
```
```{r}
sped_calcs <- sped_reg %>% 
  mutate(
    # simple diff
    sped_c_diff = sped_c_20 - sped_c_15,
    # percent change
    sped_c_prccng = ((sped_c_20 - sped_c_15) / sped_c_15 * 100) %>% round_half_up(1),
    # simple diff of percent
    sped_p_ppd = sped_p_20 - sped_p_15,
    # meets threshold 8.5% 2015
    sped15_thsh = case_when(
      sped_p_15 >= 8.5 ~ "Y",
      TRUE ~ "N"
    ),
    # meets threhhold 8.5% 2020
    sped20_thsh = case_when(
      sped_p_20 >= 8.5 ~ "Y",
      TRUE ~ "N"
    )
  )

sped_calcs %>% glimpse
```
```{r}
sped_calcs %>% 
  select(
    school_name,
    sped_p_15,
    sped15_thsh,
    sped_p_20
    , sped20_thsh
  ) %>% 
  head(20)
```
```{r}
sped_calcs %>% 
  write_rds("data-processed/02_sped_calcs.rds")
```

